---
title: "The Great American Coffee Taste Test"
subtitle: "Proposal"
author: 
  - name: "Team Name"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "A visualization of survey data collected from the great American coffee taste which was conduceted by James Hoffman and Cometeer."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
#| echo: false
library(tidyverse)
library(readr)
library(janitor)
library(here)
library(fs)
library(knitr)
```

## Dataset

```{r}
#| label: load-dataset-coffee_survey
#| message: false
#| echo: false
#| #load data from coffee_survey csv file


locale=locale(encoding="latin1") #https://stackoverflow.com/questions/71532975/trying-to-read-csv-file-says-error-in-ncharx-width-invalid-multibyte-s

coffee_survey <- read_csv("data/coffee_survey.csv", locale = readr::locale(encoding = "latin1"))

#https://forum.posit.co/t/unable-to-import-hotel-bookings-dataset-to-rstudio/166402

```

```{}
```

```{r glimpse_data}

#| label: glimpse-dataset-coffee_survey
#| message: false
#| echo: false

glimpse(coffee_survey)
```

This data set was created from a partnership between James Hoffman and Cometeer. They sent four coffee samples to over 4,000 Americans and asked them to fill out this survey which asked them which coffee they liked the best from the samples, what their regular preference is, and what they typically spend on a cup of coffee. I chose this dataset as it is a survey which includes questions on demographics that can be used to categorize participants. Participants were not chosen randomly so parametric statisitcal tests can not be used to analyze this dataset, but it may still reveal interesting correlations.

```{r}
#| echo: false

dim(coffee_survey)
```

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
