---
title: "The Great American Coffee Taste Test"
subtitle: "Proposal"
author: 
  - name: "Violeta Keifer"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "A visualization of survey data collected from the great American coffee taste which was conducted by James Hoffman and Cometeer."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
  echo: true
  message: false
---

```{r install packages and set theme}

#install packages into r
if(!require("pacman"))
  install.packages("pacman")

pacman::p_load(tidyverse, 
               glue, 
               scales,
               ggthemes,
               forcats,
               rsconnect,
               shiny,
               jpeg,
               ggpubr,
               png,
               gt,
               dplyr,
               viridis,
               patchwork,
               here,
               ggridges,
               colorspace,
               stringr,
               dsbox,
               janitor,
               openintro,
               ggrepel,
               patchwork,
               readr)

# set theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))
```

## Dataset

```{r load coff_survey dataset}
#| label: load-dataset-coffee_survey
#| #load data from coffee_survey csv file


locale=locale(encoding="latin1") #https://stackoverflow.com/questions/71532975/trying-to-read-csv-file-says-error-in-ncharx-width-invalid-multibyte-s

coffee_survey_raw <- read_csv(here::here("data/coffee_survey.csv"), locale = readr::locale(encoding = "latin1"))


```

```{r tidy column names}


coffee_survey <- coffee_survey_raw |> 
  janitor::clean_names() |> 
  # Get rid of one-hot encoding; users can do that if they'd like. Also,
  # "flavorings" columns are empty.
  dplyr::select(
    submission_id,
    age = what_is_your_age,
    cups = how_many_cups_of_coffee_do_you_typically_drink_per_day,
    where_drink = where_do_you_typically_drink_coffee,
    brew = how_do_you_brew_coffee_at_home,
    brew_other = how_else_do_you_brew_coffee_at_home,
    purchase = on_the_go_where_do_you_typically_purchase_coffee,
    purchase_other = where_else_do_you_purchase_coffee,
    favorite = what_is_your_favorite_coffee_drink,
    favorite_specify = please_specify_what_your_favorite_coffee_drink_is,
    additions = do_you_usually_add_anything_to_your_coffee,
    additions_other = what_else_do_you_add_to_your_coffee,
    dairy = what_kind_of_dairy_do_you_add,
    sweetener = what_kind_of_sugar_or_sweetener_do_you_add,
    style = before_todays_tasting_which_of_the_following_best_described_what_kind_of_coffee_you_like,
    strength = how_strong_do_you_like_your_coffee,
    roast_level = what_roast_level_of_coffee_do_you_prefer,
    caffeine = how_much_caffeine_do_you_like_in_your_coffee,
    expertise = lastly_how_would_you_rate_your_own_coffee_expertise,
    starts_with("coffee"),
    prefer_abc = between_coffee_a_coffee_b_and_coffee_c_which_did_you_prefer,
    prefer_ad = between_coffee_a_and_coffee_d_which_did_you_prefer,
    prefer_overall = lastly_what_was_your_favorite_overall_coffee,
    wfh = do_you_work_from_home_or_in_person,
    total_spend = in_total_much_money_do_you_typically_spend_on_coffee_in_a_month,
    why_drink = why_do_you_drink_coffee,
    why_drink_other = other_reason_for_drinking_coffee,
    taste = do_you_like_the_taste_of_coffee,
    know_source = do_you_know_where_your_coffee_comes_from,
    most_paid = what_is_the_most_youve_ever_paid_for_a_cup_of_coffee,
    most_willing = what_is_the_most_youd_ever_be_willing_to_pay_for_a_cup_of_coffee,
    value_cafe = do_you_feel_like_you_re_getting_good_value_for_your_money_when_you_buy_coffee_at_a_cafe,
    spent_equipment = approximately_how_much_have_you_spent_on_coffee_equipment_in_the_past_5_years,
    value_equipment = do_you_feel_like_you_re_getting_good_value_for_your_money_with_regards_to_your_coffee_equipment,
    gender,
    gender_specify = gender_please_specify,
    education_level,
    ethnicity_race,
    ethnicity_race_specify = ethnicity_race_please_specify,
    employment_status,
    number_children = number_of_children,
    political_affiliation
  )

glimpse(coffee_survey)
```

This data set was created from a partnership between James Hoffman and Cometeer. They sent four coffee samples to over 4,000 Americans and asked them to fill out this survey which asked them which coffee they liked the best from the samples, what their regular preference is, and what they typically spend on a cup of coffee. I chose this data set as it includes many possible variables to analyze, including demographics for respondents. The original survey can be viewed [here.](https://cometeer.com/pages/the-great-american-coffee-taste-test?utm_source=google&utm_medium=paid-search&utm_campaign=PMax%7CBrand%7CUS&utm_term=&utm_content=&gad_source=1&gclid=Cj0KCQjwmMayBhDuARIsAM9HM8cGTK7xXEIM7blKzNjvdJvSO3naRo3OW4lFtnvNYls0c46kFHsUQ_AaAoEREALw_wcB) This project will examine the impact of age and on participants ratings of acidity, bitterness, and preference for the four sampled coffees.

To begin exploring the data, the raw file was loaded and the data was tidied up using janitor and the suggested column titles from the Tidy Tuesday page.

The survey had 48 questions and produced a data set with 4,042 responses.

```{r dimensions, include = TRUE}

dim(coffee_survey)
```

Nearly half of the responses were from people who were aged 25 to 34 years old. A quarter of participants were aged 34 to 44 years old. The distribution of ages is skewed to the left with very few participants under 18 years old.

```{r demographic exploration, fig.width = 10}

coffee_survey |>
  filter(!is.na(age)) |>
  ggplot(aes(x = fct_relevel(age, c("<18 years old", "18-24 years old", "25-34 years old", "35-44 years old", "45-54 years old", "55-64 years old", ">65 years old")))) +
  geom_bar(fill = "brown") +
  labs(x = "Age",
       y = "Frequency",
       title = "Ages of Coffee Survey Participants")


  
```

## Question

How did bitterness, acidity, and personal preference vary for each coffee sample according to the age of participants?

## Analysis Plan

-   The first visualization will examine the impact of age on coffee preference. This will be a stacked bar chart with age on the y-axis and frequency on the x-axis. The stacked bars will each correspond to one of the four coffee samples.
-   The second visualization will show how each age group rated the acidity and bitterness of each of the four coffee samples. This will be a grouped bar chart with age on the y-axis and frequency on the x-axis. Each bar in the groups will correspond to one of the four cofee samples.
    -   An issue that will be need to be addressed is that the number of responses for each question varies as participants did not answer every question on the survey. The number of participants for each variable will need to be deteremined and participants may need to be removed if they did not answer questions related to the four variables of age, bitterness, acidity, and overall preference.
    -   Some age groups also got considerably less responses than other groups such as the \<18 years old group. This group may be excluded in the final analysis.

## Task Checker

+----------------+-------------+---------+----------+---------------------------------------------------------------------+
| Task Name      | Status      | Due     | Priority | Summary                                                             |
+:===============+:============+:========+:=========+:====================================================================+
| Upload Data    | Done        | 5/22/24 | High     | Uploaded and tidied coffee_survey                                   |
+----------------+-------------+---------+----------+---------------------------------------------------------------------+
| Explore Data   | Done        | 5/29/24 | Medium   | Explore demographics of survey                                      |
+----------------+-------------+---------+----------+---------------------------------------------------------------------+
| Proposal       | Done        | 5/29/24 | High     | Finish survey and publish site                                      |
+----------------+-------------+---------+----------+---------------------------------------------------------------------+
| Peer Proposal  | Done        | 6/3/24  | High     | Make adjustments based on peer feedback                             |
+----------------+-------------+---------+----------+---------------------------------------------------------------------+
| Final Proposal | Done        | 6/10/24 | Medium   | Make adjustments from feedback                                      |
+----------------+-------------+---------+----------+---------------------------------------------------------------------+
| Tidy Data      | In progress | 6/10/24 | Medium   | Check n for each variable                                           |
|                |             |         |          |                                                                     |
|                |             |         |          | Remove participants that did not answer all 4 questions of interest |
+----------------+-------------+---------+----------+---------------------------------------------------------------------+
| Create Graphs  | In progress | 6/17/24 | Medium   | Create all graphs and place into slides                             |
+----------------+-------------+---------+----------+---------------------------------------------------------------------+
| Presentation   | Not started | 6/24/24 | High     | Create final presentation video.                                    |
+----------------+-------------+---------+----------+---------------------------------------------------------------------+
